---
title: "Programming in Base R"
format: pdf
editor: visual
---

```{r}
library(tidyverse)
library(palmerpenguins)
```

## Task 1

### Question 1

```{r}
?read_csv
```

We cannot use read_csv specifically because our delimiter for the given data is a semicolon, not a comma. There are 2 other functions described in the help file called read_csv2 and read_delim which allow for more control over the delimiter, and read_csv2 specifically uses ; as the delimiter which we will use.

```{r}
data <- read_csv2("data/data.txt")
data
```

### Question 2

```{r}
data2 <- read_delim("data/data2.txt", delim = '6', col_types = "fdc")
data2
```

## Task 2

### Question 1

```{r}
trailblazer <- read_csv("data/trailblazer.csv")
glimpse(trailblazer)
```

### Question 2

```{r}
trailblazer_longer <- trailblazer |>
  pivot_longer(cols = 2:11, 
               names_to = c('Game', 'Location'),
               names_sep = '_',
               values_to = 'Points')
head(trailblazer_longer, 5)
```

### Question 3

```{r}
trailblazer_longer |>
  pivot_wider(names_from = Location, values_from = Points) |>
  group_by(Player) |>
  summarise(mean_home = mean(Home, na.rm = T),
            mean_away = mean(Away, na.rm = T),
            diff = mean_home - mean_away) |>
  arrange(desc(diff))
```

Jusuf Nurkic, Robert Covington and Nassir Little were the top three players who scored more points at home than away throughout the first 10 games of the season.

## Task 3

### Question 1

```{r}
?penguins
penguins |>
  select(species, island, bill_length_mm) |>
  pivot_wider(
    names_from = island, values_from = bill_length_mm
    )
```

We did not group by the species and island, so there are errors when we run this code since there are multiple values for bill_length_mm. <NULL> means that there are no penguins for that combination of species and island. \<dbl \[52\]\> means that for the Adelie species on Torgersen, there are 52 dbl values, so there are 52 different values of bill_length_mm for that group. <list> means that the column is a list since it is comprised of many dbls or null values.

### Question 2

```{r}
penguins |>
  group_by(species, island) |>
  summarise(count = n()) |>
  pivot_wider(names_from = island, values_from = count, values_fill = 0)
```

## Task 4

### Question 1

```{r}
penguins_nona <- penguins |>
  mutate(bill_length_mm = case_when(
    (species == 'Adelie') & is.na(bill_length_mm) ~ 26,
    (species == 'Gentoo') & is.na(bill_length_mm) ~ 30,
    TRUE ~ bill_length_mm
  ))

head(penguins_nona, 10)
```
